### 2019-03-06 response/response-demography-1.tex

\section*{Editor}

First, the description of the problem and literature review are sparse.
The front end of the paper should be revised to situate the study in the
wider literature on son preference and sex selection in the region. The
contributions of the current study should be highlighted. R1 offered
particularly useful ideas in this regard. These changes would make the
paper appeal to the broad readership of Demography.

[Response:]

I have completely rewritten the introduction to better situate the paper in the literature
and to highlight the contributions of the paper.

@@

- Spacing as an indicator for son preference
- The relationship between spacing and child health
- The relationship between spacing and women's health
- Important for understanding how spacing in general behaves. See last point from
editor: if lower fertility, better health, and increased female labor market participation 
are all is associated with a compression of spacing, the use of sex selection would
work against this trend and make it look like India is not behaving like other countries.
- Women's role in general

- complete rewrite of introduction
- highlight how method is different in ``Estimation Strategy'' section.

@@

[Make the paper about spacing; need to emphasize why spacing is interesting]

5. The use of competing hazard models is not new in demographic research
and doesn’t require long exposition. This section can be cut
sustainably. I don’t think the methodological innovation is a key
contribution of the paper as the methods are established even if they
have not been used to examine spacing.

[Response:]

I have reduced the length of the section by more than 1/4 (the original section was 5
pages long, while the revised section is less than 3.5).
The first two paragraphs in the section now highlight how my approach is different from
prior research, especially the non-randomness of birth outcome and non-proportionality
assumption.

There are three important differences between my approach and what is standard
in the literature, and I believe further shortening would negatively impact readers'
ability to understand the results.
First, most research that use survival models employ proportional hazard models, so 
motivating and providing more information on the non-proportional set-up is important.
Second, the use of conditional survival curves is, to my knowledge, new, and, therefore,
requires a detailed discussion.
Third, the sex ratio calculations is a new way to use competing risk estimates and is 
required to understand whether the use of sex selection vary within spells.

### 2019-03-06 paper/sex_selection_spacing-ver2.tex
Minor edits
Need better description of why my estimation method better

The unit of analysis is a spell, the period from one birth to the next.
The model is a discrete time, non-proportional, competing risk hazard model with two exit 
states: either a boy or a girl is born.
The standard approach in the spacing literature is to use a single-exit model, but with
access to sex selection the sex of the next child is not random and parents' choices 
will impact the spacing to a girl or a boy differentially.%
\footnote{
\cite{Merli2000} used a discrete hazard model to examine whether 
there were under-reporting of births in rural China, although they 
estimated separate waiting time regressions for boys and girls.
}
Both the non-randomness and differential spacing can be captured with a competing risk 
set up.

I use a non-proportional hazard specification, which allows the shape of the hazard 
functions to vary across groups, rather than the standard proportional hazard function 
for three reasons.
First, it is unlikely---even in the absence of prenatal sex determination---that the 
effect of household characteristics such as the sex composition of previous births have 
the same effect throughout the entire spell.%
\footnote{
Assuming that the effect of sex composition is the same throughout a spell is especially 
problematic for higher-order spells where different sex composition of previous births can 
lead to substantial differences both in the likelihood of progressing to the next birth and
how soon couples want their next child if they are going to have one.
}
Second, the introduction of prenatal sex determination exacerbates any bias from the 
proportionality assumption since sex-selective abortion use varies across groups,
which affects birth spacing, and because a household's use of sex selection may vary 
within a spell, which means that the effects of covariates vary within the spell as well.
Finally, the use of a non-proportional specification mitigates any potential effects 
of unobserved heterogeneity when used in conjunction with a flexible baseline hazard 
\citep{Dolton1995}.

For each woman, $i=1,\ldots,n$, the starting point for a spell is time $t=1$, and 
the spell continues until time $t_i$, when either a birth occurs or the spell 
is censored.%
\footnote{
The time of censoring is assumed independent of the hazard rate,
as is standard in the literature.
}
There are two exit states: the birth of a boy, $j=1$, or the birth of a girl, $j=2$, and 
$J_i$ is a random variable indicating which event took place.
The discrete time hazard rate $h_{ijt}$ is specified as
\begin{equation}
 h_{ijt} = \frac{\exp(D_j(t) + \alpha_{jt}'\mathbf{Z}_{it} + \beta_j'\mathbf{X}_{i})} 
 {1 + \sum_{l=1}^2 \exp(D_j(t) + \alpha_{lt}'\mathbf{Z}_{it} + \beta_l'\mathbf{X}_{i})} \: \: \; \; \;  j = 1,2
 \label{eq:hazard}
\end{equation}
where the explanatory variable vectors, $\mathbf{Z}_{it}$ and $\mathbf{X}_{i}$, capture 
individual, household, and community characteristics discussed below,
and $D_{j}(t)$ is the piece-wise linear baseline hazard for outcome $j$, captured
by dummies and the associated coefficients,
\begin{equation}
D_j(t) = \gamma_{j1} D_1 + \gamma_{j2} D_2 + \ldots + \gamma_{jT} D_T,
\end{equation}
with $D_m = 1$ if $t=m$ and zero otherwise.
This approach to modeling the baseline hazard is flexible and does not place 
overly strong restrictions on the baseline hazard.

@@

\section{ }

>>>>

The standard approach in the birth spacing literature is to use proportional hazard
models with a single exit, the birth of a child.
There are two problems with the standard approach in this setting.
First, and most importantly, the introduction of sex selection means that the sex of the 
next child is no longer necessarily random and that parents' choices will impact the 
spacing to the birth of a girl or a boy differentially.
I, therefore, use a competing risk set-up, which can capture both the non-randomness of
the birth outcome and the differential spacing to the birth of a girl or a boy.
To my knowledge, this is the first application of competing risk models to birth
spacing.%
\footnote{
\cite{Merli2000} used a discrete hazard model to examine whether 
there were under-reporting of births in rural China, although they 
estimated separate waiting time regressions for boys and girls.
}

Second, it is unlikely---even in the absence of prenatal sex determination---that the 
effect of household characteristics such as the sex composition of previous births have 
the same effect throughout the entire spell as is assumed by the proportional hazard
model.
Assuming that the effect of sex composition is the same throughout a spell is especially 
problematic for higher-order spells where different sex composition of previous births can 
lead to substantial differences both in the likelihood of progressing to the next birth 
and how soon couples want their next child if they are going to have one.
The introduction of prenatal sex determination exacerbates any bias from the 
proportionality assumption since sex-selective abortion use varies across groups,
which affects birth spacing, and because a household's use of sex selection may vary 
within a spell, which means that the effects of covariates vary within the spell as well.
I, therefore, use a non-proportional hazard specification, which allows the shape of the 
hazard functions to vary across groups.
The use of a non-proportional specification also mitigates any potential effects 
of unobserved heterogeneity when used in conjunction with a flexible baseline hazard 
\citep{Dolton1995}.

The model is a discrete time, non-proportional, competing risk hazard model with two exit 
states: either a boy or a girl is born.
The unit of analysis is a spell, the period from one birth to the next.
For each woman, $i=1,\ldots,n$, the starting point for a spell is time $t=1$, and 
the spell continues until time $t_i$, when either a birth occurs or the spell 
is censored.%
\footnote{
The time of censoring is assumed independent of the hazard rate,
as is standard in the literature.
}
There are two exit states: the birth of a boy, $j=1$, or the birth of a girl, $j=2$, and 
$J_i$ is a random variable indicating which event took place.
The discrete time hazard rate $h_{ijt}$ is specified as
\begin{equation}
 h_{ijt} = \frac{\exp(D_j(t) + \alpha_{jt}'\mathbf{Z}_{it} + \beta_j'\mathbf{X}_{i})} 
 {1 + \sum_{l=1}^2 \exp(D_j(t) + \alpha_{lt}'\mathbf{Z}_{it} + \beta_l'\mathbf{X}_{i})} \: \: \; \; \;  j = 1,2
 \label{eq:hazard}
\end{equation}
where the explanatory variable vectors, $\mathbf{Z}_{it}$ and $\mathbf{X}_{i}$, capture 
individual, household, and community characteristics discussed below,
and $D_{j}(t)$ is the piece-wise linear baseline hazard for outcome $j$, captured
by dummies and the associated coefficients,
\begin{equation}
D_j(t) = \gamma_{j1} D_1 + \gamma_{j2} D_2 + \ldots + \gamma_{jT} D_T,
\end{equation}
with $D_m = 1$ if $t=m$ and zero otherwise.
This approach to modeling the baseline hazard is flexible and does not place 
overly strong restrictions on the baseline hazard.

@@

\section{Results\label{sec:results}}
